---
title: "miniprojekt"
author: "Rasmus Kragh Jørgensen"
date: "4/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(tidyverse)
```

1. Investigate to begin with the effect of prednison by considering estimates of the survival functions for the treatment groups and by performing
a log-rank test.
```{r}
data <- read.table("data.txt",header = F)
colnames(data) <- c("patientnumber","death","obstime","treatment","sex","ascites","age","prothrombin")
data <- as.data.frame(data)



fit <- survfit(Surv(data$obstime, data$death)~data$treatment,data=data)
plot(fit,lty=c(1,2))#solid=placebo dashed=prednison
survdiff(Surv(data$obstime, data$death)~data$treatment)#log-rank test

```
We observe that the survival functions crosses each other. We observe from the log-rank test that the p-value is 0.4 wich indicates that there is no difference between the two groups. 



2. Fit a Cox proportional model with all covariates (main effects) included.
Perform a careful assessment of the fitted model (including “Andersen”
plots). Also check whether there are any problematic issues with the
covariates (use histograms, boxplots,...).

```{r}
source("andersenplot.R")
#fit.cox <- coxph(Surv(data$obstime, data$death)~strata(data$treatment+data$sex+data$ascites)+data$age+data$prothrombin,data = data)
fit.cox <- coxph(Surv(data$obstime, data$death)~strata(data$treatment)+data$sex+data$ascites+as.numeric(data$age)+data$prothrombin,data = data)
andersen.fit <- survfit(fit.cox)


fit.cox2 <- coxph(Surv(data$obstime, data$death)~strata(data$sex)+data$treatment+data$age+data$ascites+data$prothrombin,data = data)
andersen.fit2 <- survfit(fit.cox2)

fit.cox3 <- coxph(Surv(data$obstime, data$death)~strata(data$ascites)+data$treatment+data$sex+data$age+data$prothrombin,data = data)
andersen.fit3 <- survfit(fit.cox3)

par(mfrow=c(1,3))
andersen.plot(andersen.fit,main="treatment")
andersen.plot(andersen.fit2,main="sex")
andersen.plot(andersen.fit3,main="ascites")
```
It looks quite nice. in the ascites plot we obsverse a single outlier


```{r}
par(mfrow=c(1,2))
hist(data$age %>% as.numeric,xlab="age")
boxplot(data$age %>% as.numeric)
```
From the histogram we observe that the distributions is skewed to the right with a single observation to the left. 
The boxplot shows that there are some observations who are outliers.  
```{r}
par(mfrow=c(1,2))
hist(data$prothrombin ,xlab="prothrombin")
boxplot(data$prothrombin )
```
We observe from the historgram that there is a small tail to the right. The boxplot indicates that there isnt any point outside of the quantiles. 


3. Within the framework of the fitted Cox regression model, test the null
hypothesis of no effect of prednison treatment.
```{r}
fit.cox <- coxph(Surv(data$obstime, data$death)~strata(data$treatment)+data$sex+data$ascites+as.numeric(data$age)+data$prothrombin,data = data)
summary(fit.cox)


fit.coxtest <- coxph(Surv(data$obstime, data$death)~as.factor(data$treatment)+data$sex+data$ascites+data$age+data$prothrombin,data = data)
#summary(fit.coxtest)
```
From the summary, we observe that the test-p-values are very low, hence we reject the null-hopothesis that the treatment has no effect. 



4. Investigate whether some kind of transformations should be applied to
the continuous covariates age and prothrombin. Here you can use martingale residuals, discretization of the covariates, comparison of models obtained with different transformations etc.

We try to remove both age and prothrombin, one at the time
```{r}
#we remove Age
fit_2 <- coxph(Surv(data$obstime,data$death)~strata(data$treatment)+
                     data$sex+data$ascites++data$prothrombin,data = data)
#We remove prothombin
fit_3 <- coxph(Surv(data$obstime,data$death)~strata(data$treatment)+
                     data$sex+data$ascites+as.numeric(data$age),data = data)
summary(fit_2)
summary(fit_3)
mart_2 <- residuals(fit_2)
mart_3 <- residuals(fit_3)

scatter.smooth(data$age,mart_2,main="Age vs no-age")
scatter.smooth(data$prothrombin,mart_2,main="Prothrombin vs no-prothrombin")

```
Both scatter smooth plots looks approximately linear. Now we try to discretization of age and prothombin. But first we consider a histogram of age

```{r}
data$age <- data$age %>% as.numeric()
hist(data$age)
```
We discretize age in the interval 0-40,40-55,55-70,70-80

```{r}
data_trans_disc <- data
data_trans_disc$age <-data_trans_disc$age %>% cut(breaks=c(0,40,55,70,80)) 
fit_4 <- coxph(Surv(obstime,death)~strata(treatment)+sex+ascites+age,data = data_trans_disc)
summary(fit_4)
mart_4 <- residuals(fit_4)
scatter.smooth(data$age,mart_4,main="Disc of age")
```
Again it looks pretty linear. From the summery we observe that the interval age(40,55] is non-significant and another interval might be more appropiate. now we discretize prothombin.

```{r}
hist(data$prothrombin)
data_trans_disc <- data
data_trans_disc$prothrombin <-data_trans_disc$prothrombin %>% cut(breaks=c(0,60,80,140)) 
fit_5 <- coxph(Surv(obstime,death)~strata(treatment)+sex+ascites+prothrombin,data = data_trans_disc)
summary(fit_5)
mart_5 <- residuals(fit_5)
scatter.smooth(data$prothrombin,mart_5,main="Disc of prothombin")
```



log-transform of age
```{r}

data_trans <- data
data_trans$age <-data_trans$age %>% as.numeric() %>% log() 
fit.trans <- coxph(Surv(data_trans$obstime,data_trans$death)~strata(data_trans$treatment)+
                     data_trans$sex+data_trans$ascites+data_trans$age+data_trans$prothrombin,data = data_trans)
summary(fit.trans)
mart.res <- residuals(fit.trans)
scatter.smooth(data$age,mart.res)#martingale residuals

AIC(fit_2,fit_3,fit_4,fit_5)
```
We have conducted a log transformation of age effect. We observe that 



5. Investigate whether there are indications of effects of interaction between treatment and the other covariates (this is a kind of data mining/exploratory analysis).

```{r}
fit_interact <- coxph(Surv(data$obstime,data$death)~data$treatment*data$ascites+ data$sex+data$ascites+data$age+data$prothrombin,data = data)
summary(fit_interact)
AIC(fit_interact)
```
We tried to see the interaction between treatment and age, prothombin and sex, but these where non-significant, But the interaction between treatment and asciets is significant. 


6. Give an estimate for the survival function for a prednison treated man,
who at the start of the treatment was 57, had Ascites=little, and a
prothrombin index equal to 85. What is the probability that he will
survive to be 65 ?


```{r}
fit_interact <- coxph(Surv(obstime,death)~treatment*ascites+ sex+ascites+age+prothrombin,data = data)

df_new <- data.frame(treatment=1,
                     age=57,
                     ascites=1,
                     prothrombin=85,
                     sex=1)
esti_surv <- survfit(fit_interact,newdata = df_new)
esti_surv
summary(esti_surv,times = 2922)
```

We observe a survival probability of 25.5%
